---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

library(captioner)
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/captioner)](http://cran.r-project.org/package=captioner)

# Status

*captioner* is no longer under development as the functionality has been replaced
with the new package [bookdown](https://github.com/rstudio/bookdown). Thank you
to everyone who used and contributed!

# Overview

*captioner* is an R package for generating figure numbers and captions

Information about recent updates can be found in [NEWS.md](https://github.com/adletaw/captioner/blob/master/NEWS.md).

# Installation

*captioner* is now available on [CRAN](https://cran.r-project.org/web/packages/captioner/index.html).
If you would like to install the current development version:

* `install.packages("devtools")`
* `devtools::install_github("adletaw/captioner")`

or if you want the vignette:

* `devtools::install_github("adletaw/captioner", build_vignettes = TRUE)`
* `vignette("using_captioner")`

# Basic Usage

A call to the function `captioner()` returns a captioner function for each set of figures, tables, etc. that you want to create.

You can generate a full caption:

```{r ex_1a}
fig_nums <- captioner()
fig_nums("my_first_figure", "My first figure's caption.")
```

And display it under your plot using inline code chunks:

```{r ex_1b}
plot(cars)
```

`r fig_nums("my_first_figure")`

Once you have created the caption, you can display it by using your function and specifying the name of the caption you created.

```{r ex_2}
fig_nums("my_first_figure")
```

The `display` parameter allows you to adjust how much of the caption is displayed.  For example, you can also generate a figure reference:

```{r ex_3}
fig_nums("my_first_figure", display = "cite")
```

Which can be displayed inline using inline code chunks, like so (`r fig_nums("my_first_figure", display = "cite")`).

Each new call to your captioner function will create a caption with an incremented number.  For example:

```{r ex_4}
fig_nums("my_second_figure", "The caption for my second figure.")
```

If you need to create a new set of captions with new numbering, for example, for your tables, simply make a new call to the `captioner` function.  Using the `prefix` parameter allows you to specify the prefix associated with your new captions:

```{r ex_5}
table_nums <- captioner(prefix = "Table")
table_nums("cat_table", "A table of cat breeds.")
```

|Breed          |Coat       |
|---            |---        |
|Abyssinian     | Short     |
|Aegean         | Semi-Long |
|Persian        | Long      |

`r table_nums("cat_table")`

One way to make in-text citations easier is by using `partial`<sup>1</sup> from the package [pryr](https://github.com/hadley/pryr) to create a shorter version of the function.  For example:

```{r ex_6}
citef <- pryr::partial(fig_nums, display = "cite")
citef("my_second_figure")
```

Take a look at the [vignette](https://github.com/adletaw/captioner/tree/master/vignettes/using_captioner.Rmd) for many more details.

***

1.  I learned this from the glorious H Wickham and [you can too](http://adv-r.had.co.nz/)
